\section{Formal Semantics} \label{sec:model}

To articulate our approach and \lang's semantics, we present a core model
smart contract language \corelang equipped with
behavioral contracts. The language modulo behavioral
contracts is entirely a subset of Solidity.
This section presents \corelang's abstract syntax, static semantics, and
translation semantics.

\input{fig:syntax}

\subsection{Syntax}

\Cref{fig:syntax} shows the abstract syntax of \corelang,
which models the most important portion of Solidity.
\corelang's type universe contains integers, unsigned integers, booleans,
addresses, mappings, and struct.
A mapping type has the domain type and codomain type.
A struct type has a name and a list of type-and-identifier declarations.

At the top-level, a contract $c$ consists of a list of type-and-identifier
declarations and a list of function definitions.
Those type-and-identifier declarations specify storage states.
As in other languages, a function definition has a name, parameter
declarations, return types, and the function body.  To model boundaries between
multiple smart contracts, a function can be either \Keywd{public} or
\Keywd{private}.
In \corelang, a specification $\sigma$ is additionally attached to every function.

We can also define interfaces at top-level, which only contain function
declarations. Interfaces are useful when invoking functions on addresses,
providing expected types of arguments and returned values for the address call.

The body of functions is simply a statement.
Statements can be a type-and-identifier declaration (without providing an initial
value), an expression (e.g. calling a function for its side effects), a
composition of two statements, an assignment, a \Keywd{return} statement, a \Keywd{revert} statement,
or a conditional statement.
The left-hand side part of assignments (i.e. assignable) can be either
a type-and-type declaration, an identifier, or a projection of mappings or structs.
Since \corelang permits both mappings and structs, which can recursively refer to
other types, projection allows use to directly access nested data structures,
e.g. \code{v.field1[0].field2}.
A \Keywd{revert} statement aborts the execution, reverting any changes made to
the blockchain state.

Expressions could be an identifier, a literal value (e.g. numbers), a binary
operation, projections of mappings or structs, address/function calls.
Address and function calls have form $\AddrCall{e_1}{(x: e_2)^*}{e_3}$,
where $e_1$ is the target callee expression,
$(x: e_2)^*$ are call options (e.g. specifying the gas budget),
and $e_3^*$ are call arguments.
An ordinary function call, struct constructor call, or casting omits the call
option part, consisting only of the function name and argument expressions.

A specification is attached to a target, which is either a function name or
an address (expressed by an access path due to the existence of nested structures).
When specified for addresses, the programmer provides an interface name at the beginning.
In addition, a specification introduces bindings for call options,
ordinary arguments, and returned values. The following snippet shows an example
of function specification
\[
  \overbrace{\mathit{vote}}^{\text{callee}}\Keywd{\{}\underbrace{(x : y)^*}_{\text{call options}}\Keywd{\}(}\overbrace{x^*}^{\text{arguments}}\Keywd{)} ~\Keywd{returns (}\underbrace{y^*}_{\text{returned values}}\Keywd{)}
\]
The precondition and postcondition are denoted in the \code{requires}-clause and
\code{ensures}-clause, respectively.
In practice not all functions need to be specified with conditions.
In such cases, a no-op specification is expressed where the precondition and
postcondition are simply a boolean expression \code{true}.
In the \code{where}-clause, programmers can specify the condition for addresses
appeared as arguments or returned values for the parent call, recursively using
the same syntax.
When the specification is attached to an address, we allow very flexible ways
to specify the callee. For example, if a struct argument contains an address
value that we would like to guard:
\begin{lstlisting}
// x is a struct argument whose field y is an address
... where { x.y.f requires $\mathit{precond}$ ensures $\mathit{postcond}$ }
\end{lstlisting}
we can use the projection syntax to specify
the condition when address \code{x.y} is called on function \code{f}.

\subsection{Static Semantics}
\input{fig:static}

The syntax permits arbitrary expressions appeared as preconditions and
postconditions.
However, not all possible expression are valid executable specifications.
For example, a programmer can write arithmetic expressions \code{requires x + 1} as a precondition,
which do have a meaning for our purpose.
The static semantics presented here concerns about when a specification is
considered valid.

In general, the precondition and postcondition need to be well-typed, which
entails that the specified conditions are of type boolean, and only
use well-scoped variables.
Since \corelang is a subset of Solidity except for specifications, we focus on
the typing for specifications and omit the typing for the rest of the language,
referring to the informal Solidity design document.

%dynamic semantics
%\yy{do we want to model the context of contracts? (instead of reasoning one contract)}

\subsection{Translation Semantics}

\note{GW: what if in the condition, user calls a guarded function?}

\paragraph{Extent of Effective Checking}
monitors all value flows within the extent

\paragraph{Wrapping Addresses} \lang transform \Keywd{address} type to a struct
type with an additional type $\mathcal{G}$ that carries the pre-/postcondition.
At runtime when the underlying address is called,
corresponding checks can be performed by examining this value of type $\mathcal{G}$.

$$
\Keywd{address} \leadsto \Keywd{address} \times \mathcal{G}
$$

\paragraph{Unwrapping Addresses} balabala

\subsection{Soundness}

In this section, we discuss when and where \corelang provides the guarantee
that a violation will be detected (thus the execution of the smart contract
will be reverted).

Our translation semantics identifies the extent of effective checking.
Notably, if an address value is stored to the state or passed to other
contracts, the address leaves the boundary and loses its guards by unwrapping.


\corelang's semantics guarantee relatively complete monitoring, i.e.


