\documentclass[sigconf,review,anonymous]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

\setcopyright{none}
\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column

%\setcopyright{acmcopyright}
% \setcopyright{none}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[ICSE 2024]{46th International Conference on Software Engineering}{April 2024}{Lisbon, Portugal}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

% \settopmatter{printacmref=false}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.8]{beramono}
\usepackage{amsmath}
\usepackage{color}
\usepackage{xcolor,colortbl}
\usepackage{url}
\usepackage{listings}
\usepackage{paralist}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{flushend}
\usepackage{bcprules}
\usepackage{textcomp}
\usepackage{pdfpages}
\usepackage{cleveref}
\usepackage{xspace}
\usepackage{stackengine}
\usepackage{stmaryrd}
\usepackage{multirow}
\usepackage{xurl}

\definecolor{light-gray}{gray}{0.85}
\input{macros}

\begin{document}

%\title{Contracts for Contracts \\
%  {\Large Securing Smart Contracts with Higher-Order Temporal Behavioral Contracts}
%}

\title{Consolidating Smart Contracts with Behavioral Contracts}

\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \country{USA}
  \postcode{43017-6221}
}

\iffalse
\lstMakeShortInline[keywordstyle=,%
  flexiblecolumns=false,%
  %basewidth={0.56em, 0.52em},%
  mathescape=false,%
  basicstyle=\ttfamily\small]@
\fi

\begin{abstract}
  Ensuring the reliability of smart contracts is of vital importance.
  However, existing static and dynamic verification tools are not expressive 
  or convenient enough for many rich properties.
  Moreover, as part of the programming tool, they do not offer programmers a notion 
  to articulate the interface between components.
  We argue that the ``design-by-contract'' approach should complement the development
  of smart contract programs. Unfortunately, there is only minimal linguistic support
  in existing smart contract languages.

  In this paper, we introduce a Solidity language extension \lang that supports
  behavioral contracts. \lang offers programmers a modular specification
  and monitor system for both first-order and higher-order behaviors. % (e.g. functions taking addresses as arguments).
  %We describe \lang with examples, static semantics, and translation semantics.
  We evaluate \lang using \numCaseStudied real-world cases, showing 
  the effectiveness of \lang 
  % that our approach is effective 
  in expressing critical conditions and preventing attacks.
  % Using 33 cases, w
  We also evaluate \lang's efficiency and compare the gas consumption with manually inserted assertions,
  % their gas consumption of \lang  compared to manually inserted assertions,
  showing that our approach only introduces minimal gas overhead.
  By separating specifications and implementations, \lang helps programmers 
  write not only more robust but also more readable code.
  
  %\lang offers an appealing approach to building robust smart contract
  %programs without sacrificing readability and maintainability.
  
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\iffalse
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}
\fi

%\keywords{smart contracts, behavioral contracts, specification, runtime verification}

\maketitle

\input{01_introduction}
\input{02_background}
\input{03_contract_solidity}
\input{04_formal_model}
\input{05_implementation}
\input{06_case_study}
\input{07_evaluation}
\input{08_related_work}
\input{09_conclusion}

%\section*{Acknowledgement}
\newpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{10_references}

\end{document}

% Local Variables:
% TeX-engine: luatex
% TeX-master: t
% End:
